---
title: "Adapting reintroduction tactics in successive trials increases the likelihood of establishment for an endangered carnivore in a fenced sanctuary"
author: "Wilson B A, Evans M J, Batson W G, Banks S C, Gordon I J, Fletcher D B, Wimpenny C, Newport J, Belton E, Rypalski A, Portas T & Manning A D"
date: "2 August 2023"
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: true
    theme: cerulean
    highlight: pygments
editor_options: 
  chunk_output_type: console
  
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'tutorial.html')) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```

# **Data preparation**

First, we installed the [pacman Package Management Tool](https://cran.r-project.org/web/packages/pacman/index.html), which allows us to install and load subsequent packages in a condensed and efficient way. 

```{r, eval=FALSE}
#install.packages("pacman")
```

```{r, results='hide', warning=FALSE, message=FALSE}
# Install and load required packages
pacman::p_load(adehabitatLT, brglm, boot, effects, ggmap, ggplot2, 
               ggpubr, janitor, lme4, lsmeans, maptools, multcomp, 
               MuMIn, plyr, readr, readxl, rgdal, rstudioapi, sp, 
               tidyverse, viridis)
```

We also set the working directly to where this R markdown is saved using the `rstudioapi` package.

```{r}
# Set the working directory to where this markdown is saved
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

# **Calculating analysis**

```{r, results='hide', warning=FALSE, message=FALSE}
# Assign raw data filename to an object
raw_data <- "data.xlsx"

# Read in den location data
den <- read_excel(raw_data, sheet="den locations") %>% 
  clean_names() %>% 
  rename(easting=x, northing=y)

# Read in and project MFWS fence shapefile using rgdal
mfws <- readOGR(dsn="shapefiles/mfws_fence.shp", verbose=FALSE) %>% 
  spTransform(CRS("+proj=utm +zone=55 +ellps=WGS84")) %>%
  # Transform shapefile into a dataframe
  fortify(verbose=FALSE) %>%
  mutate(lat=as.numeric(lat + 10000000), 
         long=as.numeric(long))

# Plot den locations on MFWS map for each trial
ggplot() + 
  geom_path(mfws, mapping=aes(x=long, y=lat, group=group), col="grey15") +
  geom_jitter(den, mapping=aes(x=easting, y=northing, 
                               col=factor(year_of_acquisition)), alpha=0.2) +
  coord_sf(xlim=c(695750, 699250), ylim=c(6104250, 6107750)) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_rect(fill="white"),
        axis.text.x = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        legend.key = element_blank(), 
        strip.background = element_blank()) +
  facet_wrap(~factor(year_of_acquisition), ncol=3) +
  scale_colour_manual(values = viridis(3, begin=0.9, end=0.1), name="Fence") +
  xlab("") + ylab("") + labs(col="Trial")
```

## Cluster dens

  1. Create clusters for dens within 10 m of one another

```{r}
den <- den %>% 
  # Remove some founders
  subset(bank_id!="EQ15" & bank_id!="EQ16" & bank_id!="EQ17" & 
           # Remove duplicated rows
           den_id!="119" & den_id!="1682" & 
           # Select only animals that survived the establishment period
           history=="Alive" & 
           # Remove rows with no valid coordinates
           easting!="NA" & 
           # Select only rows within 42 days of release
           days_post_release<43)

# Create an xy dataframe
xy <- data.frame(cbind(den$easting, den$northing))

# Create a distance matrix
distxy <- dist(xy)

# Hierarchically cluster the distance matrix
chc <- hclust(distxy)

# Calculate distance between dens with a 10 m threshold
chc.d10 <- cutree(chc, h=12)

# Append cluster labels to the den df
den$cluster <- chc.d10
```

  2. Calculate mean eastings and northings for each cluster

```{r}
meanclusters <- plyr::ddply(den, .(cluster), summarize,  
                            x2=mean(easting), y2=mean(northing))
vlookup <- merge(den, meanclusters, by= "cluster" )

# Sort df by den_id
vlookup <- arrange(vlookup, den_id)

# Sort df by den_id
den <- arrange(den, den_id)

den$mean_x <- vlookup$x2
den$mean_y <- vlookup$y2
```

## Distance per day

Calculate distances travelled for the first quoll, "EQ01".

```{r}
# Subset to consecutive days only
den <- subset(den, days_between==1)

# Remove irrelevant animal
eq <- subset(den, bank_id=="EQ01")

# Convert column to factor
eq$bank_id <- factor(eq$bank_id)

# Convert date to a form R can interact with
pdate <- as.POSIXct(strptime(as.character(eq$posi_xdate), "%Y.%m.%d ")) 
data_xy=eq[c("mean_x", "mean_y")]

# Create a Spatial Points class for all locations
xysp <- SpatialPoints(data_xy) 
proj4string(xysp) <- CRS("+proj=utm +zone=55 +ellps=WGS84")

# Create a spatial df of coordinates
sppt <- data.frame(xysp) 

# Create a spatial df of identities
idsp <- data.frame(eq$bank_id)

# Merge identities and dates into same spatial df
merge <- data.frame(idsp) 

# Merge identities and date to coordinates
coordinates(merge) <- sppt 

# Create an object of class ltraj to store movements
move <- as.ltraj(xy=eq[,c("mean_x", "mean_y")], date=pdate, id=idsp) 

# Plot movements and save to jpeg file
jpeg(file="EQ01_plot.jpeg", width=5500, height=5000, units="px", res=800)
  # Plot movements
  plot(move, xlim=c(695750, 699250), ylim=c(6104250, 6107750), 
       xlab="Easting", ylab="Northing", main="EQ01")
  plot(mfws, add=TRUE)
dev.off()

dist_eq <- move[[1]]
dist_eq$bank_id <- "EQ01"
sum.table <- cbind(eq, dist_eq)
```

```{r}
dist_eq <- move[[1]]
dist_eq$bank_id <- "EQ01"
sum.table <- cbind(eq, dist_eq)

# Create a vector the loop looks at to get values
eq_ids <- levels(den$bank_id) 

# Remove the first identity to avoid duplicating later
eq_ids <- eq_ids[c(2:length(eq_ids))] 

for(i in eq_ids){
  eq <- subset(den, bank_id==i)
  eq$bank_id <- factor(eq$bank_id) #converts to factor
  pdate <- as.POSIXct(strptime(as.character(eq$posix_date),"%Y.%m.%d")) #convert date to a format POSIX needs
  data_xy=eq[c("mean_x", "mean_y")]
  xysp <- SpatialPoints(data_xy) #creates a class Spatial Points for all locations
  proj4string(xysp) <- CRS("+proj=utm +zone=55 +ellps=WGS84")
  
  sppt <- data.frame(xysp) #creates a Spatial Data Frame from sppt
  idsp <- data.frame(eq$bank_id) #creates a spatial data frame of ID
  merge <- data.frame(idsp) #merges ID and Date into the same spatial data frame
  coordinates(merge) <- sppt #adds ID and Date data frame with locations data frame
  move <- as.ltraj(xy=eq[,c("mean_x","mean_y")], date=pdate, id=idsp) #creates an object of class ltraj to store movements
  
  jpeg(file=paste(i, "plot.jpeg", sep="_"), 
       width=5500, height=5000, units="px", res=800)
  plot(move, xlim=c(695750, 699250), ylim=c(6104250, 6107750), 
       xlab="Easting", ylab="Northing", main=paste(i))
  plot(mfws, add=T)
  dev.off()
  
  dist_eq <- move[[1]]
  dist_eq$bank_id <- i
  eqtable <- cbind(eq, dist_eq)
  sum.table <- rbind(sum.table, eqtable)
}
```

## Distance from release site

```{r}
den$dfr <- sqrt((den$release_x-den$mean_x)^2 + 
                  (den$release_y-den$mean_y)^2)
```

# **Modelling**

## Data preparation

```{r}
# Read in identities data
animal <- read_excel(raw_data, sheet="identities") %>% 
  clean_names() %>% 
  mutate(year = factor(year), 
         moved_perc = as.numeric(as.character(moved_perc)), 
         distance_mean = as.numeric(as.character(distance_mean)), 
         distance_mean_zero = as.numeric(as.character(distance_mean_zero)), 
         # Create binary survival probability variable
         fate = ifelse(fate=="Survived", 1, 0)) %>%
  subset(bank_id!="EQ15" & bank_id!="EQ16" & bank_id!="EQ17" & 
           bank_id!="EQ25" & bank_id!="EQ32" & bank_id!="EQ33" & 
           bank_id!="EQ34" & bank_id!="EQ35" & bank_id!="EQ36" & 
           bank_id!="EQ37" & bank_id!="EQ38" & bank_id!="EQ39" & 
           year!="2017" & year!="2018" & year!="2018")

femanimal <- subset(animal, sex!="M" & year!="2016")

codenanimal <- subset(animal, bank_id!="EQ02" & bank_id!="EQ05" & 
                        bank_id!="EQ10" & bank_id!="EQ11" & 
                        bank_id!="EQ12" & bank_id!="EQ29")
```

## Survival

### By trial (fig_ 2A, model 1)

```{r}
animal$trial <- factor(animal$trial)
mod <- glm(fate ~ trial, data=animal, 
           family=binomial(link=logit))
summary(mod)
anova(mod, test=c("Chisq"))

pw <- glht(mod, mcp(trial="Tukey"))  
ph <- cld(pw, alpha=0.050, Letters=letters, adjust="tukey")
summary(glht(mod, mcp(trial="Tukey")))

jpeg(file=paste("Survival by trial (fig_ 2A, model 1).jpeg"), 
     width=2500, height=2500, units="px", res=800)
effects <- allEffects(mod)
effects <- data.frame(effects$trial)
effects$trial <- factor(effects$trial)
effects$tukey <- c("a", "b", "b")

mod1 <- ggplot(effects, aes(x=trial, y=fit, group=1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), 
                colour="orange", width=0.1) +
  geom_point(shape=21, size=3, 
             fill="orange", colour="orange") +
  theme(axis.line=element_line(colour="black"), legend.position="none",
        panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
        panel.background=element_blank(), axis.title=element_text(),
        axis.text.y=element_text(angle=0, vjust=0.5,color="black"),
        axis.text.x=element_text(angle=0, vjust=0.5,color="black"),
        strip.text.y=element_blank()) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), 
                     labels=seq(0, 1, 0.2), 
                     limits=c(0, 1.04)) +
  xlab("Trial") + ylab("Probability of survival") +
  geom_text(aes(x=trial, y=upper, label=tukey), nudge_y=0.05)
dev.off()
mod1
```

### By origin (model 2)

```{r}
mod <- glm(fate ~ origin, data=animal, 
           family=binomial(link=logit))
summary(mod)
anova(mod, test=c("Chisq"))

    mod <- glm(fate ~ origin, data=animal20162017, 
               family=binomial(link=logit))
    summary(mod)
    anova(mod, test=c("Chisq"))

pw <- glht(mod, mcp(origin="Tukey"))  
ph <- cld(pw, alpha=0.050, Letters=letters, adjust="tukey")
summary(glht(mod, mcp(origin="Tukey")))

jpeg(file=paste("Survival by origin (model 2).jpeg"), 
     width=2500, height=2500, units="px", res=800)
effects <- allEffects(mod)
effects <- data.frame(effects$origin)
effects$origin <- factor(effects$origin)
effects$tukey <- "a"

mod2 <- ggplot(effects, aes(x=origin, y=fit, group=1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1) +
  geom_point(shape=21, size=3, fill="black", colour="black") +
  theme(axis.line=element_line(colour="black"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.background=element_blank(),
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"),
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"),
        axis.title=element_text(),
        strip.text.y=element_blank()) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), 
                     labels=seq(0, 1, 0.2), 
                     limits=c(0, 1.1)) + 
  xlab("Origin") + ylab("Probability of survival") +
  geom_text(aes(x=origin, y=upper, label=tukey), nudge_y=0.06)
dev.off()
```

### By sex (fig 2B, model 3)

```{r}
animal$sex <- factor(animal$sex)
mod <- glm(fate ~ sex, data=animal, 
           family=binomial(link=logit))
summary(mod)
anova(mod, test=c("Chisq"))

    animal2016$sex <- factor(animal2016$sex)
    mod <- glm(fate ~ sex, data=animal2016, 
               family=binomial(link=logit))
    summary(mod)
    anova(mod, test=c("Chisq"))
    
pw <- glht(mod, mcp(sex="Tukey"))  
ph <- cld(pw, alpha=0.050, Letters=letters, adjust="tukey")
summary(glht(mod, mcp(sex="Tukey")))

jpeg(file=paste("By sex (fig 2B, model 3).jpeg"), 
     width=2500, height=2500, units="px", res=800)
effects <- allEffects(mod)
effects <- data.frame(effects$sex)
effects$sex <- factor(effects$sex)
effects$tukey <- c("a", "b")

mod3 <- ggplot(effects, aes(x=sex, y=fit, group=1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), color="brown", width=0.1) +
  geom_point(shape=21,size=3, fill="brown", color="brown") +
  theme(axis.line=element_line(color="black"), 
        legend.position="none", 
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.background=element_blank(),
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"), 
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"), 
        axis.title=element_text(),
        strip.text.y=element_blank()) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), 
                     labels=seq(0, 1, 0.2), 
                     limits=c(0, 1.04)) + 
  xlab("Sex") + ylab("Probability of survival") +
  geom_text(aes(x=sex, y=upper, label=tukey), nudge_y=0.05)

dev.off()
mod3
```

### By den sharing (model 4)

```{r}
mod <- glm(fate ~ coden_pa, data=codenanimal, 
           family=binomial(link=logit))
summary(mod)
anova(mod, test=c("Chisq"))

pw <- glht(mod, mcp(coden_pa="Tukey"))  
ph <- cld(pw, alpha=0.050, Letters=letters, adjust="tukey")
summary(glht(mod, mcp(coden_pa="Tukey")))

jpeg(file=paste("Survival by den sharing (model 4).jpeg"), 
     width=2500, height=2500, units="px", res=800)
effects <- allEffects(mod)
effects <- data.frame(effects$coden_pa)
effects$tukey <- c("a", "a")

mod4 <- ggplot(effects, aes(x=coden_pa, y=fit, group=1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1) +
  geom_point(shape=21, size=3, fill="black", colour="black") + 
  theme(axis.line=element_line(colour="black"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.background=element_blank(),
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"),
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"),
        axis.title=element_text(),
        strip.text.y=element_blank()) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), 
                     labels=seq(0, 1, 0.2), 
                     limits=c(0, 1)) +
  xlab("") + ylab("Probability of survival") +
  geom_text(aes(x=coden_pa, y=upper, label=tukey), nudge_y=0.06)
dev.off()
```

### By pouch young (model 5)

```{r}
mod <- glm(fate ~ pypa, data=femanimal, 
           family=binomial(link=logit))
summary(mod)
anova(mod, test=c("Chisq"))

pw <- glht(mod, mcp(pypa="Tukey"))  
ph <- cld(pw, alpha=0.050, Letters=letters, adjust="tukey")
summary(glht(mod, mcp(pypa="Tukey")))

jpeg(file=paste("Survival by pouch young (model 5).jpeg"), 
     width=2500, height=2500, units="px", res=800)
effects <- allEffects(mod)
effects <- data.frame(effects$pypa)
effects$pypa <- factor(effects$pypa)
effects$tukey <- c("a", "a")

mod5 <- ggplot(effects, aes(x=pypa, y=fit, group=1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1) +
  geom_point(shape=21, size=3, fill="black", colour="black") +
  theme(axis.line=element_line(colour="black"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.background=element_blank(),
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"),
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"),
        axis.title=element_text(),
        strip.text.y=element_blank()) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), 
                     labels=seq(0, 1, 0.2), 
                     limits=c(0, 1)) +
  xlab("Pouch young") + ylab("Probability of survival") +
  geom_text(aes(x=pypa, y=upper, label=tukey), nudge_y=0.06) #not working...

dev.off()
```

## Den sharing

### By trial (model 6)

```{r}
codenanimal$trial <- factor(codenanimal$trial)
mod <- glm(coden_pc ~ trial, data=codenanimal)
summary(mod)
anova(mod, test=c("Chisq"))

pw <- glht(mod, mcp(trial="Tukey"))  
ph <- cld(pw, alpha=0.050, Letters=letters, adjust="tukey")
summary(glht(mod, mcp(trial="Tukey")))

#this plot is not right
jpeg(file=paste("Den sharing by trial (model 6).jpeg"), 
     width=2500, height=2500, units="px", res=800)
effects <- allEffects(mod)
effects <- data.frame(effects$trial)
effects$Year <- factor(effects$trial)
effects$tukey <- c("a", "a", "a")

mod6 <- ggplot(effects, aes(x=trial, y=fit, group=1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1) +
  geom_point(shape=21, size=3, fill="black", colour="black") +
  theme(axis.line=element_line(colour="black"),
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        panel.background=element_blank(), 
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"),
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"),
        axis.title=element_text(), 
        strip.text.y=element_blank()) + 
  xlab("Trial") + ylab("Percentage of fixes \n found den sharing (%)") +
  geom_text(aes(x=trial, y=upper, label=tukey), nudge_y=3) #letters not in right position...

dev.off()
```

### By sex (model 7)

```{r}
mod <- glm(coden_pc ~ sex, data=codenanimal)
summary(mod)
anova(mod, test=c("Chisq"))

pw <- glht(mod, mcp(sex="Tukey"))  
ph <- cld(pw, alpha=0.050, Letters=letters, adjust="tukey")
summary(glht(mod, mcp(sex="Tukey")))

#this plot is not right
jpeg(file=paste("Den sharing by sex (model 7).jpeg"), 
     width=2500, height=2500, units="px", res=800)
effects <- allEffects(mod)
effects <- data.frame(effects$sex)
effects$sex <- factor(effects$sex)
effects$tukey <- "a"

mod7 <- ggplot(effects, aes(x=sex, y=fit, group=1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1) +
  geom_point(shape=21, size=3, fill="black", colour="black") +
  theme(axis.line=element_line(colour="black"),
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        panel.background=element_blank(), 
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"),
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"),
        axis.title=element_text(),
        strip.text.y=element_blank()) +
  xlab("Sex") + ylab("Percentage of fixes \n found den sharing (%)") +
  geom_text(aes(x=sex, y=upper, label=tukey), nudge_y=3)
dev.off()
```

### By origin (model 7.5)

```{r}
mod <- glm(CodenPC ~ origin, data=codenanimal)
summary(mod)
anova(mod,test=c("Chisq"))

pw <- glht(mod, mcp(origin="Tukey"))  
ph <- cld(pw, alpha=0.050, Letters=letters, adjust="tukey")
summary(glht(mod, mcp(origin="Tukey")))

#this plot is not right
jpeg(file=paste("Den sharing by origin (model 7).jpeg"), 
     width=2500, height=2500, units="px", res=800)
effects <- allEffects(mod)
effects <- data.frame(effects$origin)
effects$origin <- factor(effects$origin)
effects$tukey <- "a"

mod7 <- ggplot(effects, aes(x=origin, y=fit, group=1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1) + 
  geom_point(shape=21, size=3, fill="black", colour="black") + 
  theme(axis.line=element_line(colour="black"), 
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        panel.background=element_blank(), 
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"), 
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"), 
        axis.title=element_text(), 
        strip.text.y=element_blank()) + 
  xlab("Origin") + ylab("Percentage of fixes \n found den sharing (%)") + 
  geom_text(aes(x=origin, y=upper, label=tukey), nudge_y=3)

dev.off()
```

## Movement

### Survival by movement (fig 2C, model 8)

```{r}
subdata <- animal[!is.na(animal$moved_perc),] #changed it to a subdata
subdata$fate <- as.numeric(as.character(subdata$fate)) # for some reason it was classified as a character vector
mod <- glm(fate ~ moved_perc, data=subdata, 
           family=binomial(link=logit)) #should be a binomial model
summary(mod)
anova(mod, test=c("Chisq"))

new.dat <- data.frame(moved_perc=seq(0.12, 0.857, length.out=100)) #creates a vector for the prediction function to predict for
predicted <- predict(mod, newdata=new.dat, se.fit=TRUE, type="link") #predicting on link scale
predicted <- data.frame(predicted)
plotdat <- cbind(new.dat,predicted)
plotdat$fit_r <- inv.logit(plotdat$fit) #creating the predicted value by transforming from the link scale to the response scale
plotdat$lower_r <- inv.logit(plotdat$fit-1.96*plotdat$se.fit) #creating rough 95%CIs before transforming to the response scale
plotdat$upper_r <- inv.logit(plotdat$fit+1.96*plotdat$se.fit) #there are more complicated and robust ways such as profiling and permutations...

jpeg(file=paste("Survival by movement (fig 2C, model 8).jpeg"), 
     width=2500, height=2500, units="px", res=800)

mod8 <- ggplot(plotdat, aes(x=moved_perc, ylim(0,1), y=fit_r, group=1)) + 
  geom_ribbon(aes(ymin=lower_r, ymax=upper_r, linetype=NA), 
              alpha=0.2, fill="forest green") +
  geom_line(linetype=1, colour="forest green") +
  geom_line(aes(y= upper_r ), linetype=3, 
            colour="forest green", lwd=0.5) + 
  geom_line(aes(y= lower_r), linetype=3, 
            colour="forest green", lwd=0.5) +
  theme(axis.line=element_line(colour="black"), 
        legend.position="none",
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        panel.background=element_blank(), 
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"), 
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"), 
        axis.title=element_text(), 
        strip.text.y=element_blank()) + 
  scale_x_continuous(breaks=seq(0, 1, 0.2), 
                     labels=seq(0, 1, 0.2), 
                     limits=c(0, 1.04)) + 
  scale_y_continuous(breaks=seq(0, 1, 0.2), 
                     labels=seq(0, 1, 0.2), 
                     limits=c(0, 1.04)) + 
  xlab("Proportion of days moved") + ylab("Probability of survival")

dev.off()
mod8
```

### Survival by distance (model 9)

```{r}
subdata <- animal[!is.na(animal$distance_mean),] #changed it to a subdata
subdata$fate <- as.numeric(as.character(subdata$fate)) # for some reason it was classified as a character vector

mod <- glm(fate ~ distance_mean, data=subdata, 
           family=binomial(link=logit)) #model including zeros
summary(mod)
anova(mod, test=c("Chisq"))

new.dat <- data.frame(distance_mean=seq(180, 1050, length.out=100)) #creates a vector for the prediction function to predict for
predicted <- predict(mod, newdata=new.dat, se.fit=TRUE, type="link") #predicting on link scale
predicted <- data.frame(predicted)
plotdat <- cbind(new.dat,predicted)
plotdat$fit_r <- inv.logit(plotdat$fit) #creating the predicted value by transforming from the link scale to the response scale
plotdat$lower_r <- inv.logit(plotdat$fit-1.96*plotdat$se.fit) # creating rough 95%CIs before transforming to the response scale
plotdat$upper_r <- inv.logit(plotdat$fit+1.96*plotdat$se.fit)

jpeg(file=paste("Survival by distance (model 9).jpeg"), 
     width=2500, height=2500, units="px", res=800)

mod9 <- ggplot(plotdat, aes(x=distance_mean, y=fit_r, group=1)) + 
  geom_ribbon(aes(ymin=lower_r, ymax=upper_r, linetype=NA), 
              alpha=0.2, colour="grey40", fill="grey40") + 
  geom_line(linetype=1,colour="grey40") + 
  geom_line(aes(y= upper_r ),linetype=3, colour="grey40", lwd=0.5) + 
  geom_line(aes(y= lower_r),linetype=3, colour="grey40", lwd=0.5) + 
  theme(axis.line=element_line(colour="black"), 
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        panel.background=element_blank(), 
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"),
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"),
        axis.title=element_text(), 
        strip.text.y=element_blank()) + 
  scale_y_continuous(breaks=seq(0, 1, 0.2), 
                     labels=seq(0, 1, 0.2), 
                     limits=c(0, 1)) + 
  xlab("Distance moved between dens (m)") + ylab("Probability of survival")

dev.off()
```

**Excluding zeros**

```{r}
subdata <- animal[!is.na(animal$distance_mean_zero),] #changed it to a subdata
subdata$fate <- as.numeric(as.character(subdata$fate)) # for some reason it was classified as a character vector

mod <- glm(fate ~ distance_mean_zero, data=subdata, 
           family=binomial(link=logit)) #model including zeros
summary(mod)
anova(mod,test=c("Chisq"))

mod <- glm(fate ~ distance_mean_zero, data=subdata, 
           family=binomial(link=logit)) #model excluding zeros
summary(mod)
anova(mod, test=c("Chisq"))

new.dat <- data.frame(distance_mean_zero=seq(180, 1050, length.out=100)) #creates a vector for the prediction function to predict for
predicted <- predict(mod, newdata=new.dat, se.fit=TRUE, type="link") #predicting on link scale
predicted <- data.frame(predicted)
plotdat <- cbind(new.dat,predicted)
plotdat$fit_r <- inv.logit(plotdat$fit) #creating the predicted value by transforming from the link scale to the response scale
plotdat$lower_r <- inv.logit(plotdat$fit-1.96*plotdat$se.fit) # creating rough 95%CIs before transforming to the response scale
plotdat$upper_r <- inv.logit(plotdat$fit+1.96*plotdat$se.fit)

jpeg(file=paste("Survival by distance (nozeros, model 9).jpeg"), 
     width=2500, height=2500, units="px", res=800)

mod9 <- ggplot(plotdat, aes(x=distance_mean_zero, y=fit_r, group=1)) + 
  geom_ribbon(aes(ymin=lower_r, ymax=upper_r, linetype=NA), 
              alpha=0.2, colour="grey40", fill="grey40") + 
  geom_line(linetype=1, colour="grey40") + 
  geom_line(aes(y= upper_r ), linetype=3, 
            colour="grey40", lwd=0.5) + 
  geom_line(aes(y= lower_r), linetype=3, 
            colour="grey40", lwd=0.5) + 
  theme(axis.line=element_line(colour="black"), 
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        panel.background=element_blank(), 
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"), 
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"), 
        axis.title=element_text(), 
        strip.text.y=element_blank()) + 
  scale_y_continuous(breaks=seq(0, 1, 0.2), 
                     labels=seq(0, 1, 0.2), 
                     limits=c(0, 1)) + 
  xlab("Distance moved between dens (m)") + ylab("Probability of survival")

dev.off()
```

### By trial (fig 3A, model 10)

```{r}
codenanimal$trial <- factor(codenanimal$trial)

mod <- glm(moved_perc ~ trial, data=codenanimal)
summary(mod)
anova(mod, test=c("Chisq"))

pw <- glht(mod, mcp(trial="Tukey"))  
ph <- cld(pw, alpha=0.050, Letters=letters, adjust="tukey")
summary(glht(mod, mcp(trial="Tukey")))

jpeg(file=paste("Movement by trial (fig 3A, model 10).jpeg"), 
     width=2500, height=2500, units="px", res=800)
effects <- allEffects(mod)
effects <- data.frame(effects$trial)
effects$trial <- factor(effects$trial)
effects$tukey <- c("a", "b", "b")

mod10 <- ggplot(effects, aes(x=trial, y=fit, group=1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), 
                colour="orange", width=0.1) + 
  geom_point(shape=21, size=3, 
             fill="orange", colour="orange") + 
  theme(axis.line=element_line(colour="black"), 
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        panel.background=element_blank(), 
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"), 
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"), 
        axis.title=element_text(), 
        strip.text.y=element_blank()) + 
  scale_y_continuous(breaks=seq(0, 1, 0.2), 
                     labels=seq(0, 1, 0.2), 
                     limits=c(0, 1.04)) + 
  xlab("Trial") + ylab("Proportion of days moved") +
  geom_text(aes(x=trial, y=upper, label=tukey), nudge_y=0.05) 
dev.off()
mod10
```

### By origin (model 11)

```{r}
mod <- glm(moved_perc ~ origin, data=animal)
summary(mod)
anova(mod, test=c("Chisq"))

    mod <- glm(moved_perc ~ origin, data=animal20162017)
    summary(mod)
    anova(mod, test=c("Chisq"))

pw <- glht(mod, mcp(origin="Tukey"))  
ph <- cld(pw, alpha=0.050, Letters=letters, adjust="tukey")
summary(glht(mod, mcp(origin="Tukey")))

jpeg(file=paste("Movement by origin (model 11).jpeg"), 
     width=2500, height=2500, units="px", res=800)
effects <- allEffects(mod)
effects <- data.frame(effects$origin)
effects$origin <- factor(effects$origin)
effects$tukey <- "a"

mod11 <- ggplot(effects, aes(x=origin, y=fit, group=1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1) + 
  geom_point(shape=21, size=3, fill="black", colour="black") + 
  theme(axis.line=element_line(colour="black"), 
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        panel.background=element_blank(), 
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"), 
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"), 
        axis.title=element_text(), 
        strip.text.y=element_blank()) + 
  scale_y_continuous(breaks=seq(0, 1, 0.2), 
                     labels=seq(0, 1, 0.2), 
                     limits=c(0, 1)) + 
  xlab("Origin") + ylab("Proportion of days moved") + 
  geom_text(aes(x=origin, y=upper, label=tukey), nudge_y=0.06)
dev.off()
```

### By sex (fig 3B, model 12)

```{r}
mod <- glm(moved_perc ~ sex, data=animal)
summary(mod)
anova(mod, test=c("Chisq"))

    mod <- glm(moved_perc ~ sex, data=animal2016)
    summary(mod)
    anova(mod, test=c("Chisq"))

pw <- glht(mod, mcp(sex="Tukey"))  
ph <- cld(pw, alpha=0.050, Letters=letters, adjust="tukey")
summary(glht(mod, mcp(sex="Tukey")))

jpeg(file=paste("Movement by sex (fig 3B, model 12).jpeg"), 
     width=2500, height=2500, units="px", res=800)
effects <- allEffects(mod)
effects <- data.frame(effects$sex)
effects$sex <- factor(effects$sex)
effects$tukey <- c("a", "b")

mod12 <- ggplot(effects, aes(x=sex, y=fit, group=1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), colour="brown", width=0.1) + 
  geom_point(shape=21,size=3, fill="brown", colour="brown") + 
  theme(axis.line=element_line(colour="black"), 
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        panel.background=element_blank(), 
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"), 
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"), 
        axis.title=element_text(), 
        strip.text.y=element_blank()) + 
  scale_y_continuous(breaks=seq(0, 1, 0.2), 
                     labels=seq(0, 1, 0.2), 
                     limits=c(0, 1.04)) + 
  xlab("Sex") + ylab("Proportion of days moved") +
  geom_text(aes(x=sex, y=upper, label=tukey), nudge_y=0.05)
dev.off()
mod12
```

### By den sharing (fig 3C, model 13)

```{r}
hist(logit(codenanimal$moved_perc))
mod <- glm(moved_perc ~ coden_pa, data=codenanimal)
summary(mod)
anova(mod, test=c("Chisq"))

pw <- glht(mod, mcp(coden_pa="Tukey"))  
ph <- cld(pw, alpha=0.050, Letters=letters, adjust="tukey")
summary(glht(mod, mcp(coden_pa="Tukey")))

jpeg(file=paste("Movement by den sharing (fig 3C, model 13).jpeg"), 
     width=2500, height=2500, units="px", res=800)
effects <- allEffects(mod)
effects <- data.frame(effects$coden_pa)
effects$tukey <- c("a", "b")

mod13 <- ggplot(effects, aes(x=coden_pa, y=fit, group=1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), 
                colour="forest green", width=0.1) + 
  geom_point(shape=21, size=3, fill="forest green", colour="forest green") + 
  theme(axis.line=element_line(colour="black"), 
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        panel.background=element_blank(), 
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"),
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"),
        axis.title=element_text(), 
        strip.text.y=element_blank()) + 
  scale_y_continuous(breaks=seq(0, 1, 0.2), 
                     labels=seq(0, 1, 0.2), 
                     limits=c(0, 1.04)) + 
  xlab("") + ylab("Proportion of days moved") + 
  geom_text(aes(x=coden_pa, y=upper, label=tukey), nudge_y=0.05)

dev.off()
```

### By pouch young (model 14)

```{r}
mod <- glm(moved_perc ~ pypa, data=femanimal)
summary(mod)
anova(mod, test=c("Chisq"))

pw <- glht(mod, mcp(pypa="Tukey"))  
ph <- cld(pw, alpha=0.050, Letters=letters, adjust="tukey")
summary(glht(mod, mcp(pypa="Tukey")))

jpeg(file=paste("Movement by pouch young (model 14).jpeg"), 
     width=2500, height=2500, units="px", res=800)
effects <- allEffects(mod)
effects <- data.frame(effects$pypa)
effects$pypa <- factor(effects$pypa)
effects$tukey <- c("a", "a")

mod14 <- ggplot(effects, aes(x=pypa, y=fit, group=1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1) +
  geom_point(shape=21,size=3,fill="black",colour="black") +
  theme(axis.line=element_line(colour="black"),
        #legend.position="none" #legend
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.background=element_blank(),
        axis.text.y=element_text(angle=0, vjust=0.5, color="black"), #y axis tick label
        axis.text.x=element_text(angle=0, vjust=0.5, color="black"),
        axis.title=element_text(),
        strip.text.y=element_blank()) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), 
                     labels=seq(0, 1, 0.2), 
                     limits=c(0, 1)) +
  xlab("Pouch young") + ylab("Probability of survival") +
  geom_text(aes(x=pypa, y=upper, label=tukey), nudge_y=0.06) #not working...

dev.off()
```

## Combine plots

```{r}
fig2 <- ggarrange(mod1, mod3 + rremove("ylab") + 
                    rremove ("y.text") + rremove ("y.ticks"), 
                  mod8 + rremove("ylab") + 
                    rremove ("y.text") + rremove ("y.ticks"), 
                  heights=c(3,3),widths=c(3,3), #y axis titles
                label.x=0.89, label.y=0.99, hjust=-0.5, vjust=1.5, #figure labels
                #labels=c("a","b", "c"), #figure labels
                ncol=3, nrow=1, #align="h", #arrangement of matrix
                font.label=list(size=10, face="bold", color ="black"))

print(fig2)

ggsave(filename="fig_2.tiff", fig2, width=190, height=75, units="mm")


fig3 <- ggarrange(mod10, mod12 + rremove("ylab") + 
                    rremove ("y.text") + rremove ("y.ticks"), 
                  mod13 + rremove("ylab") + 
                    rremove ("y.text") + rremove ("y.ticks"), 
                  heights=c(3, 3),widths=c(3,3) , #y axis titles
                label.x=0.89, label.y=0.99, hjust=-0.5, vjust=1.5, #figure labels
                #labels=c("a","b","c"), #figure labels
                ncol=3, nrow=1, #align="h", #arrangement of matrix
                font.label=list(size=10, face="bold", color ="black"))

print(fig3)

ggsave(filename="fig_3.tiff", fig3, width=190, height=75, units="mm")
```

# **Model selection**

## Survival by trial and sex

```{r}
animal <- animal %>% 
  mutate(trial = factor(trial), 
         sex = factor(sex))

mod1 <- glm(fate ~ trial + sex, data=animal, 
            family=binomial(link=logit))

mod2 <- glm(fate ~ trial, data=animal, 
            family=binomial(link=logit))

mod3 <- glm(fate ~ sex, data=animal, 
            family=binomial(link=logit))

AICc(mod1, mod2, mod3)
summary(mod2)
```

## Movement by trial and sex

```{r}
mod1 <- glm(moved_perc ~ trial + sex, data=animal)
mod2 <- glm(moved_perc ~ trial * sex, data=animal)
mod3 <- glm(moved_perc ~ trial, data=animal)
mod4 <- glm(moved_perc ~ sex, data=animal)

AICc(mod1, mod2, mod3, mod4)
summary(mod2)
```

## Combine plots

```{r}
fig2 <- ggarrange(mod1, mod3 + rremove("ylab") + 
                    rremove ("y.text") + rremove ("y.ticks"), 
                  mod8 + rremove("ylab") + 
                    rremove ("y.text") + rremove ("y.ticks"), 
                  heights=c(3, 3),widths=c(3, 3), #y axis titles
                label.x=0.89, label.y=0.99, hjust=-0.5, vjust=1.5, #figure labels
                labels=c("a", "b", "c"), #figure labels
                ncol=3, nrow=1, #align="h", #arrangement of matrix
                font.label=list(size=10, face="bold", color ="black"))

# Display plot
print(fig2)

ggsave(filename="figure 2.tiff", fig_2, 
       width=190, height=75, units="mm")

fig3 <- ggarrange(mod10, mod12 + rremove("ylab") + 
                    rremove ("y.text") + rremove ("y.ticks"), 
                  mod13 + rremove("ylab") + 
                    rremove ("y.text") + rremove ("y.ticks"), 
                  heights=c(3, 3), widths=c(3,3) , #y axis titles
                label.x=0.89, label.y=0.99, hjust=-0.5, vjust=1.5, #figure labels
                labels=c("a", "b", "c"), #figure labels
                ncol=3, nrow=1, #align="h", #arrangement of matrix
                font.label=list(size=10, face="bold", color ="black"))
# Display plot
print(fig3)

ggsave(filename="figure 3.tiff", fig3, 
       width=190, height=75, units="mm")
```

# **Session info**

```{r}
# Display version information for R, OS, and packages
sessionInfo()
```